Optimal Algorithms for Dynamic Balanced Graph Partitioning

### 2020-05-14 paper/disc20clustering.tex
correcting ground set def

We provide a~lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm for the learning problem.
Later we elaborate on how to efficiently transform it to a~lower bound for the general partitioning problem.
The lower bound requires $k\geq 3$.
In contrast, for $k=2$ the learning problem is trivial: immediate collocation of communicating pairs is $1$-competitive.
(A partitioning problem for $k=2$ is non-trivial, a~lower bound of $3$ is known~\cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)

Across this paper, we often refer to groups of communicating nodes.
We use this concept slightly differently in the lower bound and the upper bounds.
In our algorithms, we group nodes with a~communication history into \emph{components}, and in the lower bound, we group nodes that may communicate into \emph{ground sets}.

A~\emph{ground set} is a group of nodes that repeatedly communicate if they are not collocated.
If an algorithm splits any pair of nodes from a ground set, then the adversary continues to produce requests to this pair until the algorithm collocates it \maciek{learning model: does not need to continue issuing requests.}
Any competitive algorithm must maintain a~perfect partition of ground sets into clusters.
We emphasize that ground sets are revealed only upon their nodes are split.
The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
Once we construct a ground set, it lasts until the end of the input sequence.
We say that a ground set is \emph{singleton} if it contains exactly one node.
We say that the node that belongs to a singleton ground set is \emph{isolated}.

>>>>

We provide a~lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm for the learning problem.
Later, we elaborate on how to efficiently transform it to a~lower bound for the general partitioning problem.
The lower bound requires $k\geq 3$.
In contrast, for $k=2$ the learning problem is trivial: immediate collocation of communicating pairs is $1$-competitive.
In contrast,
the general partitioning problem for $k=2$ is non-trivial (see Section~\ref{sec:k2}).
%, a~lower bound of $3$ is known~\cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)

Across this paper, we often refer to groups of communicating nodes.
We use this concept slightly differently in the lower bound than the upper bounds.
In our algorithms, we group nodes with a~communication history into \emph{components}.
In this section,
we group nodes that may ever communicate, into \emph{ground sets}.

Given a perfect partition,
every subset of nodes that belong to the same cluster in this partition
is a~\emph{ground set}. 
On each inter-cluster request a ground set is revealed.
Any competitive algorithm under the learning model,
 maintains a~perfect partition of ground sets into clusters.
 An algorithm recovers the (hidden) perfect partition gradually over inter-cluster requests,
  by merging pairs of ground sets involved in these requests.

The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
Once we construct a ground set, it lasts until the end of the input sequence.
We say that a ground set is \emph{singleton} if it contains exactly one node,
that is an \emph{isolated} node.
